---
title: "NYPD Shooting Report"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  editor_options: 
    chunk_output_type: inline
---

### I have included all my commands that lead to everything as code blocks. I showed only the output of the most integral ones. I hope you enjoy :D

## Beginnings

```{r}
library(tidyverse)
library(RcppRoll)
library(lubridate)

```

The URL our data is located in

```{r}
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"
```

Loading our data into R

```{r}
nyc_data <- read_csv(url_in)
```

Changing the date column to be an actual datetime object

```{r}
nyc_data <- nyc_data %>% mutate(OCCUR_DATE = mdy(OCCUR_DATE))
```

Remove unnecessary columns that don't help our analysis

```{r}
nyc_data <- nyc_data %>% select(-c(X_COORD_CD,Y_COORD_CD,Latitude, Longitude,Lon_Lat))

nyc_data

```


Quick Summary of the data

```{r}
summary(nyc_data)
```

## Visualisation

Let's do some visualisation, for that we will first create a couple of interesting tables

-   The number of records by Boro-\>Precint
-   The total victim race
-   The total perpetrator race

```{r}
nyc_boro <- nyc_data %>% group_by(BORO,PRECINCT) %>% count()


nyc_vic_total <- nyc_data %>% group_by(VIC_RACE) %>% count()

nyc_vic_total %>% ggplot(aes(y=VIC_RACE,x=n)) + geom_bar(stat="identity") + labs(title = "Victim Race",x="Amount", y="Race")

nyc_perp_total <- nyc_data %>% group_by(PERP_RACE) %>% count()

nyc_perp_total %>% ggplot(aes(y=PERP_RACE,x=n)) + geom_bar(stat="identity") + labs(title = "Perp Race",x="Amount", y="Race")

```

How many incidents per Date? Lets also plot this

```{r}
nyc_date_total <- nyc_data %>% group_by(OCCUR_DATE) %>% count()


nyc_date_total %>% ggplot(aes(x=OCCUR_DATE,y=n)) + geom_bar(stat="identity") + labs(title = "Cases over Time",x="Date", y="Number of cases")
```

Preparing for Analysis of total cases over time and plotting for total cases over time

```{r}
nyc_date_total$total <- cumsum(nyc_date_total$n)
nyc_date_total %>% ggplot(aes(x=OCCUR_DATE,y=total)) + geom_line(stat="identity") + labs(title = "total Cases over Time",x="Date", y="Total Number of cases")
```

## Modeling

```{r}
mod <- lm( total ~ OCCUR_DATE , data=nyc_date_total)

mod2 <- lm( n ~ OCCUR_DATE, data=nyc_date_total)

nyc_date_total$pred1 <- predict(mod)

nyc_date_total$pred2 <- predict(mod2)

ggplot(nyc_date_total, aes(x=OCCUR_DATE)) +  geom_line(aes(y=total), color="black") + geom_line(aes(y=pred1), color="red") + labs(title = "total Cases over Time",x="Date", y="Total Number of cases")

```

So we can see overall, it trends downwards. Meaning the cases per day are falling steadily.

### About Bias

This Data could have some certain racial bias. Not every crime gets reported. Some neighbourhoods with racial bias might be profiled more. Or in general, the people might be misidentified.
